---
title: "Assignment - 4"
author: "Dushyant Goel"
date: "2024-10-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)

```

# Tidy data and iteration

## 1.1 Missing data and iteration
### Q1

Tidyverse library, specifically package `purrr`

### Q2
```{r}
impute_by_median <- function(x) {
  me <- median(x, na.rm=TRUE)
  impute_f <- function(z) {
    if(is.na(z)){
      return(me)
    } else {
      return(z)
    }
  }
  return(map_dbl(x, impute_f))
}

v <- c(1, 2, NA, 4)
impute_by_median(v)
```

### Q3

```{r}
x <- seq(from = 0, to = 10, by = 0.1)
y <- 5 * x + 1
df_xy <- data.frame(x, y)

df_xy %>%
  head(5)
```

### Q4
```{r}
sometimes_missing <- function(index, value) {
  if (index %% 5 == 0) {
    return(NA)
  } else {
    return(value)
  }
}
```

```{r}
df_xy_missing <- df_xy %>%
  mutate(y = map2_dbl(row_number(), x, sometimes_missing))

head(df_xy_missing, 10)
```

### Q5
```{r}
df_xy_imputed <- df_xy_missing %>%  
  mutate(y = impute_by_median(y))

head(df_xy_imputed, 10)
```
## 1.2 Tidying data with pivot functions

```{r}
folder_path <- "./"
file_name <- "HockeyLeague.xlsx"
file_path <- paste(folder_path, file_name, sep="")
wins_data_frame <- read_excel(file_path, sheet="Wins")

wins_data_frame %>%
  select(1:5) %>%
  head(3)
```

The data is not tidy. The number of wins and losses and total number of games are within the same column. In context of data analysis, it is difficult to extract and manipulate in the given formate. Numeric data should have been presented one per cell.


### Q1

```{r}

wins_tidy <- wins_data_frame %>% 
  pivot_longer(cols = -c(1), names_to= "Year", values_to = "Wins of Total") %>%
  separate(col = "Wins of Total", into = c("Wins", "Total"), sep=" of ") %>%
  rename("Teams" = 1) %>%
  mutate(Wins = as.integer(Wins)) %>%
  mutate(Total = as.integer(Total))

wins_tidy %>%
  dim()

wins_tidy %>%
  head(5)
```

```{r}
laundry <- function(sheet) {
  read_excel(file_path, sheet) %>%
  pivot_longer(cols = -c(1), names_to= "Year", values_to = "of Total") %>%
  separate(col = "of Total", into = c(sheet, "Total"), sep=" of ") %>%
  rename("Teams" = 1) %>%
  mutate(Total = as.integer(Total)) %>%
  mutate({{sheet}} := as.integer(.data[[sheet]]))
    
}

sheet = "Losses"
losses_tidy <- laundry(sheet)

losses_tidy %>% dim()
losses_tidy %>% head(5)

```

### Q3
```{r}
hockey_df <-
inner_join(wins_tidy, losses_tidy) %>%
  mutate(Draws = Total - (Wins + Losses)) %>%
  mutate(across(c(Wins, Losses, Draws), ~ . / Total, .names = "{.col}_rt"))
```

### Q4
```{r}
hockey_summary <-
hockey_df %>%
  select(-Wins, -Draws, -Losses) %>%
  group_by(Teams) %>%
  summarise(across(starts_with(c("Wins","Losses","Draws")),
             list(md=median,mn=mean),
             .names="{substring(.col,1,1)}_{.fn}")) %>% 
  arrange(desc(W_md))
  
```


## 1.3 Simulation experiments of probabilities

### Q1
```{r}
num_red_balls<-3 
num_blue_balls<-7 
total_draws<-22 
prob_red_spheres<-function(z){
total_balls<-num_red_balls+num_blue_balls 
log_prob<-log(choose(total_draws,z))+z*log(num_red_balls/total_balls)+(total_draws-z)*log(num_blue_balls /total_balls)
return(exp(log_prob)) }
num_trials<-1000 # set the number of trials set.seed(0) # set the random seed
num_reds_in_simulation <- data.frame(trial=1:num_trials) %>%
  mutate(sample_balls = map(.x=trial, ~sample(10,22, replace = TRUE))) %>%
  mutate(num_reds = map_dbl( .x=sample_balls, ~sum(.x<=3) ) ) %>% 
  pull(num_reds)
prob_by_num_reds <- data.frame(num_reds=seq(22)) %>% mutate(TheoreticalProbability=prob_red_spheres(num_reds)) %>% mutate(EstimatedProbability=map_dbl(.x=num_reds, ~sum(num_reds_in_simulation==.x))/num_trials)
```

### Q2
(Reading exercise)


## 2. Conditional probability, Bayes rule and independence

### Q1
$P(A) = 0.9$  
$P(B|A) = 0.8$
$P(B^c|A^c)=0.75$
$P(A|B) = \frac{P(A).P(B|A)}{P(B)}$  

Now, \
$P(B) = P(B|A).P(A) + P(B|A^c).P(A^c)$
$P(B) = P(B|A).P(A) + [1 - P(B^c|A^c)].[1-P(A^c)]$  
Substituting values  
$P(B) = 0.745$   

Substituting,  
$P(A|B) = 0.996$


